% -*- coding: utf-8 -*-

\input macros

%\beginchapter Appendix J. Joining the\\\TeX\ Community
\beginchapter Appendix J. 加入社团

This appendix is about grouping of another kind: \TeX\ users from around the
world have banded together to form the \TeX\ Users Group (TUG), in order
to exchange information about common problems and solutions.

A newsletter/journal called {\sl TUGboat\/} has been published
since 1980, featuring articles about all aspects of \TeX\ and \MF\null.
^^{METAFONT}
TUG has a network of ``site coordinators'' who serve as focal points of
communication for people with the same computer configurations.
Occasional short courses are given in order
to provide concentrated training in special topics; videotapes of
these courses are available for rental.
Meetings of the entire TUG membership are held at least once a year.
You can buy \TeX\ T-shirts at these meetings.

Information about membership in TUG and subscription to {\sl TUGboat\/}
is available from
\begintt
{\obeylines
\TeX\ Users Group
P.O. Box 9506
Providence RI 02940\kern.05em-9506, USA.
}
\endtt
Don't delay, write today! That number again is

\smallskip
{\obeylines
\TeX\ Users Group
P.O. Box 9506
Providence RI 02940\kern.05em-9506, USA.
}


\endchapter

[The printer] should refuse to employ wandering men,
foreigners who, after having committed some grievous error,
can easily disappear and return to their own country.
\author HIERONYMUS ^{HORNSCHUCH}, '$O\rho\mkern1mu\theta o\mkern1mu %
  \tau\upsilon\pi o\gamma\mkern-1mu %
  \rho\alpha\phi\acute\iota\alpha\varsigma$ (1608)

^^|\upsilon| ^^|\varsigma|
\bigskip

An author writing an article for publication in TUGboat
is encouraged to create it on a computer file and submit it on magnetic tape.
\author BARBARA ^{BEETON}, {\sl {\tt\bslash title} How to Prepare a File\/%
   {\tt\bslash cr}\phantom{\rm\enskip(1981)}\break %
   For Publication in TUGboat\/{\tt\bslash cr}} (1981) % TUGboat 2,1 p53

%\eject
\eject\byebye
