% -*- coding: utf-8 -*-

\input macros

%\beginchapter Chapter 1. The Name of\\the Game
\beginchapter Chapter 1. 此名有诗意

\origpageno=1

%\pageno=1 % This is page number 1, number 1,
%English words like `technology' stem from a Greek root beginning with
%the letters $\tau\epsilon\chi\ldots\,$; and this same Greek word means {\sl
%art\/} as well as technology. Hence the name \TeX, which is an
%uppercase form of $\tau\epsilon\chi$.^^{TeX (actually \TeX), meaning of}
%^^|\tau|^^|\epsilon|^^|\chi|
\pageno=1\1 % This is page number 1, number 1,
英语单词``technology''来源于以字母 $\tau\epsilon\chi\ldots\,$ 开头的希腊词根；%
并且这个希腊单词除了 technology 的意思外也有 art 的意思。%
因此，名称 \TeX\ 是 $\tau\epsilon\chi$ 的大写格式。

%Insiders pronounce the $\chi$ of \TeX\ as a Greek chi, not as an `x', so that
%\TeX\ rhymes with the word blecchhh. It's the `ch' sound in Scottish words
%like {\sl loch\/} or German words like {\sl ach\/}; it's a Spanish `j' and a
%Russian `kh'. When you say it correctly to your computer, the terminal
%may become slightly moist.
在发音时，\TeX\ 的 $\chi$ 的发音与希腊的 chi 一样，而不是``x'', %
所以 \TeX\ 与 blecchhh 押韵。%
``ch''听起来象苏格兰单词中的 {\sl loch\/} 或者德语单词中的 {\sl ach\/}; %
它在西班牙语中是``j'', 在俄语中是``kh''。%
当你对着计算机正确读出时，终端屏幕上可能有点雾。

%The purpose of this pronunciation exercise is to remind you that \TeX\ is
%primarily concerned with high-quality technical manuscripts: Its emphasis is
%on art and technology, as in the underlying Greek word. If you merely want
%to produce a passably good document---something acceptable and basically
%readable but not really beautiful---a simpler system will usually suffice.
%With \TeX\ the goal is to produce the {\sl finest\/} quality; this requires
%more attention to detail, but you will not find it much harder to go the
%extra distance, and you'll be able to take special pride in the finished
%product. ^^{beauty} % since truth is indexed, ...
这个发音练习是提醒你，\TeX\ 主要处理的是高质量的专业书稿：%
它的重点在艺术和专业方面，就象希腊单词的含义一样。%
如果你仅仅想得到一个过得去——可读下去但不那么漂亮——的文书，%
那么简单的系统一般就够用了。%
使用 \TeX\ 的目的是得到{\KT{10}最好的}质量；%
这就要在细节上花功夫，但是你不会认为它难到哪里去，并且你会为所完成的作品感到%
特别骄傲。

%On the other hand, it's important to notice another thing about \TeX's name:
%The `E' is out of kilter. This ^^{logo}
%displaced `E' is a reminder that \TeX\ is about typesetting, and it
%distinguishes \TeX\ from other system names. In fact, ^{TEX} (pronounced
%{\sl tecks\/}) is the admirable {\sl Text EXecutive\/} processor developed by
%^{Honeywell Information Systems}. Since these two system names are
%^^{Bemer, Robert, see TEX, ASCII}
%pronounced quite differently, they should also be spelled differently. The
%correct way to refer to \TeX\ in a computer file, or when using some other
%medium that doesn't allow lowering of the `E', is to type `^|TeX|'. Then
%there will be no confusion with similar names, and people will be
%primed to pronounce everything properly.
另一方面重要的是要注意到与 \TeX\ 名称有关的另一件事：%
``E''是错位的。%
这个偏移``E''的标识提醒人们，\TeX\ 与排版有关，并且把 \TeX\ 从其它系统的名称%
区别开来。%
实际上，TEX~(读音为 {\sl tecks\/}) 是 Honeywell Information Systems 的%
极好的{\sl Text EXecutive\/}处理器。%
因为这两个系统的名称读音差别很大，所以它们的拼写也不同。%
在计算机中表明 \TeX\ 文件的正确方法，或者当所用的方式无法降低``E''时，%
就要写作``TeX''。%
这样，就与类似的名称不会产生混淆，并且为人们可以正确发音提供了条件。

%\exercise After you have mastered the material in this book, what will
%you be: a \TeX pert, or a \TeX nician?
%\answer A \TeX nician (underpaid); sometimes also called a \TeX acker.
\exercise 当你已经掌握本书中的知识后，你属于 \TeX\ 专家还是 \TeX\ 爱好者？
\answer \TeX\ 爱好者（酬不抵劳）；有时称为 \TeX\ 黑客。

\endchapter

They do certainly give
very strange and new-fangled names to diseases.
\author ^{PLATO}, {\sl The Republic}, Book 3 (c.\ 375 B.C.) % 405c

\bigskip

Technique! The very word is like the shriek
Of outraged Art. It is the idiot name
Given to effort by those who are too weak,
Too weary, or too dull to play the game.
\author LEONARD ^{BACON}, {\sl Sophia Trenton\/} (1920) % composed at Stanford

\vfill\eject\byebye
